\propertitle

### 2019-07-08 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-07T14:24:05+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\mynote{Say something more about advantages of such question/answer: for
  example we can make statements about total activity of brain region even
  across recordings when sampled neurons aren't the same.}

@@

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the region are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a region, comprising a
network of $N$ neurons, from which other sets of neurons had or will have
the same probability of being recorded, even in other experimental
sessions, as the set of $n$ neurons that was actually recorded in this
session. This is our \enquote{full network}\mynote{how about calling it
  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

@@

In the neuroscientific application above, this means for example that if
pairwise correlations are sufficient statistics for a particular sample of
a network, then they \emph{cannot} be sufficient for the full networ, or
for a different sample; and vice versa, if we assume that they are
sufficient statistics for the full network, then we must expect them
\emph{not} to be sufficient for any recorded sample thereof. But questions
about sufficient statistics are implicitly meant to apply to a whole brain
area, not just to some casually recorded sample. Thus, maximum-entropy
methods applied at the sample level paradoxically give an answer opposite
to the one we might think they're giving!

The comparisons of different sufficient statistics are only approximate,
however: we ask only whether our uncertainties conditional on either are
\emph{appreciably} different in some specified sense. It's therefore worth
checking whether the different results we get from a sufficient statistics
for the sample and for the full-network are still very similar when we
consider an altogether different statistics.

@@

in-depth discussion about underlying assumptions, derivation from
probability calculus, higher-order approximations beyond maxent and
"error bars" around the distributions

%plot_mc_6mom_unif_afss3tr_N10000_t400_L500_T5e+05_s30000_a0_i3

@@

Finally, we would like to point out the usefulness and importance of the
probability formulae that relate our states of knowledge about a network
and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
formulae is essential in neuroscience, where we try to understand
properties of extended brain regions from partial observations. The
formulae presented here reflect a simple, symmetric state of ignorance.
More work is needed \citep[\cf][]{levinaetal2017} to extend these formulae
to account for finer knowledge of the cerebral cortex and its network
properties.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-08T13:40:41+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\mynote{Say something more about advantages of such question/answer: for
  example we can make statements about total activity of brain area even
  across recordings when sampled neurons aren't the same.}

@@

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain areas
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the area are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a area, comprising a
network of $N$ neurons, from which other sets of neurons had or will have
the same probability of being recorded, even in other experimental
sessions, as the set of $n$ neurons that was actually recorded in this
session. This is our \enquote{full network}\mynote{how about calling it
  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

@@

In the neuroscientific application above, this means that if pairwise
correlations, for example, are sufficient statistics for a particular
sample of from a brain area, then they \emph{cannot} be sufficient for the
full network of neurons in that area, or for a different sample; vice
versa, if we assume that they are sufficient statistics for the full
network, then we must expect them \emph{not} to be sufficient for any
recorded sample thereof. But questions about sufficient statistics are
meant to be addressed to a whole brain area, not just to some casually
recorded sample. Thus, maximum-entropy methods applied at the sample level
paradoxically give an answer opposite to the one we might think they're
giving.

Questions about sufficient statistics are therefore correctly addressed by
applying the maximum-entropy method at the full-network level. If a
comparison with the measured frequencies of the sample activities is
desired, then the marginal distribution to the sample size should be used.

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
%comparison_ME_sample_full_levels.nb
Does the application to the full-network lead to very different results
from the one to the sample? Figure~\ref{fig:comparisons_marginals} gives a
graphical answer in the case of the first data set. The upper plot shows
the maximum-entropy distributions for a full network of $10\,000$ neurons,
constructed from two moments (\textcolor{mygreen}{dashed green curve}) and
from four moments (\textcolor{myblue}{solid blue curve}). The lower plot
shows the maximum-entropy distributions for the sample (dashed curves),
from two moments (\textcolor{mygreen}{empty green triangles}) and from four
moments (\textcolor{mygreen}{empty blue squares}), and the distributions
obtained by marginalizing the full-network distribution to the sample size
(solid curves), from two moments (\textcolor{mygreen}{filled green
  triangles}) and from four moments (\textcolor{mygreen}{filled blue
  squares}). The measured frequency distribution (\textcolor{myred}{red
  circles}) is also shown. We note the following:
\begin{itemize}
\item The application to the full-network (upper plot) leads to two
  completely different distributions (even different number of modes);
  clearly the two statistics are not even approximately equivalent for
  inferential purposes.
\item The application to the sample leads to deceivingly similar
  distributions (empty triangles and empty squares), which can only be
  distinguished on a logarithmic scale. This could lead to the erroneous
  conclusion that the two statistics are approximately equivalent.
\item The difference between marginals from the full-network distribution
  and the distributions obtained at the sample level (filled vs empty
  markers) is larger than the difference between different statistics
  (triangles vs squares).
\item The marginals of the application to the full network are closer to
  the measured frequencies than the distributions obtained at the sample
  level (although this closeness is not a valid criterion for their
  goodness).
\end{itemize}
Therefore, the two maximum-entropy applications lead to different results,
and the application at the full-network level is clearly superior, leading
to a clearer distinction between the effect of considering different
statistics.

@@

in-depth discussion about underlying assumptions, derivation from
probability calculus, higher-order approximations beyond maxent and
"error bars" around the distributions
\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/MC_plot1000.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/MC_samples1000.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
%mctests3.nb
%plot_mcfd_6mom_unif_afssd26_cok_tr_N1000_t50_L50_T0.333333333333333_s60000_a0

@@

Finally, we would like to point out the usefulness and importance of the
probability formulae that relate our states of knowledge about a network
and its samples, presented in \sect~\ref{sec:prob_samples}. This kind of
formulae is essential in neuroscience, where we try to understand
properties of extended brain areas from partial observations. The
formulae presented here reflect a simple, symmetric state of ignorance.
More work is needed \citep[\cf][]{levinaetal2017} to extend these formulae
to account for finer knowledge of the cerebral cortex and its network
properties.
